### 2008-09-24 doc/a-rex/arex_tech_doc.tex
Some readability improvements, standard appearance

%% LyX 1.4.5.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside,english]{article}
\usepackage{pslatex}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\usepackage{array}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage{setspace}
\onehalfspacing
\usepackage{amssymb}
\IfFileExists{url.sty}{\usepackage{url}}
                      {\newcommand{\url}{\texttt}}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\newenvironment{LyXParagraphLeftIndent}[1]%
{
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \addtolength{\leftmargin}{#1}
    \setlength{\parsep}{0pt plus 1pt}%
  }
  \item[]
}
{\end{list}}
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newenvironment{lyxlist}[1]
{\begin{list}{}
{\settowidth{\labelwidth}{#1}
 \setlength{\leftmargin}{\labelwidth}
 \addtolength{\leftmargin}{\labelsep}
 \renewcommand{\makelabel}[1]{##1\hfil}}}
{\end{list}}
\newenvironment{lyxcode}
{\begin{list}{}{
\setlength{\rightmargin}{\leftmargin}
\setlength{\listparindent}{0pt}% needed for AMS classes
\raggedright
\setlength{\itemsep}{0pt}
\setlength{\parsep}{0pt}
\normalfont\ttfamily}%
 \item[]}
{\end{list}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{multirow}
\usepackage{verbatim}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {A.Konstantinov},
  pdftitle = {The NorduGrid Grid Manager and GridFTP Server.},
  pdfsubject = {},
  pdfkeywords = {},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%

\usepackage{babel}
\makeatother
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

\begin{tabular}{rl}

\resizebox*{3cm}{!}{\includegraphics{ng-logo}}

&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}

\end{tabular}

@@

@@

\begin{center}
\textsc{\large The ARC Computational Job Management Component - A-REX}
\par\end{center}{\large \par}

\vspace*{0.5cm}

\begin{center}
\textit{\large Description and Administrator's Manual}
\par\end{center}{\large \par}

\vspace*{1.5cm}

{\centering \large A.Konstantinov\footnote{aleks@fys.uio.no} \large \par}

\end{titlepage}

\tableofcontents{}

\newpage

\section{Introduction\label{sec:intro}}

A-REX is one of ARC middleware components that implements functions
of so called \emph{Computing Element} (CE).

It takes care of job pre- and post-processing. This comprises the
stage-in of files containing input data or program modules from wide
range of sources and the transfer or storing of output results.

The A-REX implements a Web Service (WS) interface which provides a
way to submit and control computational tasks (jobs) to be executed
by the A-REX and he underlying Local Resource Management System.

\textbf{You should use this document for advanced configuration purposes
and understanding of internals of the aforementioned tools. For general
installation and configuration of whole system refer to other documents
available at \url{http://www.nordugrid.org/papers.html}. }

@@

Each job gets an identifier (\textit{jobid}). This is a handle - WS-Addressing
\cite{ws-addr-soap} XML document - which identifies the job in the
A-REX and the Information Interface.

Jobs are initiated and controlled through the WS interface. Complete
job descriptions (JD)are passed to the A-REX through WS in JSDL\cite{jsdl}
coded description. Input data files and job executables are transferred
separately through the same interface, as described in the next section.

@@

One of the most important tasks of the A-REX is to take care of processing
input and output data (files) of the job. Input files are gathered
in the SD or associated cache area. There are two ways to put a file
into the SD:

\begin{itemize}
\item Download is initiated by the A-REX -- This is the case for files defined
in the JD (with name and source). The A-REX alone is responsible to
ensure that all required files will be available in the SD.\\
The supported protocols for sources at the moment are (in case of
full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
some less standard sources are supported. Those are described below.
The A-REX fully relies on HED framework \cite{hed} for data transferring
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-REX. Examples of such
files are the main executable of the job, the job's input files, etc.
These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and
any other remote data transport during execution of a job is not guaranteed
(at least not by A-REX).

@@

\section{\label{Section: Job Flow}Job flow}

From the point of view of the A-REX a job passes through various states.
Picture \ref{job states diagram} presents a diagram of the possible
states of a job.%
\begin{figure}[tbh]
\begin{centering}
\includegraphics{pic1}
\par\end{centering}

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the dedicated Local
Information Description Interface of A-REX using the UI or any other
suitable tool or through query method of WS interface.

@@

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to a
CE but is not processed yet. The A-REX will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The A-REX is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If all files
are successfully gathered the job moves to the next state. If \textbf{any}
file can't be downloaded or it takes the UI too long to upload a file,
the job moves to \textbf{Finishing} state. It is possible to put a
limit on the number of simultaneous \textbf{Preparing} jobs. If this
limit is exceeded, jobs ready to enter the \textbf{Preparing} state
 will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
mark. Exceptions are jobs which have no files to be downloaded. These
are processed out of limits.
\item \textbf{Submiting} - The job is being passed for execution to the
\emph{Local Resource Management System} (LRMS). The corresponding
backends for many LRMSs are provided with the default installation.
If the local job submission is successful the job moves to the \textbf{Executing}
state. Otherwise it moves to \textbf{Finishing}. It is possible to
limit the aggregate number of jobs in \textbf{Submiting} and \textbf{Executing}
states.
\item \textbf{Executing} - The job is queued or being executed in the LRMS.
The A-REX takes no actions except waiting until the job finishes.
\item \textbf{Killing} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at an Indexing Service. The user can download data files
from the SD by using the UI or other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the A-REX.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the Information Interface of the A-REX. If a
job was moved to \textbf{Finished} because of failure, it may be restarted
on request of a client. When restarted, a job is moved to the state
previous to the one in which it failed and is assigned mark PENDING.
This is needed in order to not break the configuration limits. Exception
is a job failed in \textbf{Executing} state and lacking input files
specified in JD. Such a job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by
the user}. No files will be moved to their destination SE.

\section{URLs}

In full installation the A-REX and it's components support the following
data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
http, https, lfc} and \emph{rls.} For more information please see
{}``The Hosting Environment of the Advanced Resource Connector middleware''
document \cite{hed}.

\section{\label{section: internals}Internals}

@@

For each local UNIX user listed in the A-REX configuration - including
a generic one which covers all local user identities - a \textit{control
directory} exists. In this directory the A-REX stores information
about jobs belonging to that user. Multiple users can share the same
\textit{control directory}. In most common configuration case A-REX
serves all users defined by Operating System and stores their control
file in same directory. To make it easier to recover in case of failure,
the A-REX stores most information in files rather than in memory.
All files belonging to the same job have names starting with \textbf{job.ID.,}
where ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.status} - current state of the job. This is a plain
text file containing a single word representing the internal name
of current state of the job. Possible values and corresponding external
job states are :

\begin{itemize}
\item ACCEPTED
\item PREPARING
\item SUBMIT
\item INLRMS
\item FINISHING
\item FINISHED
\item CANCELING
\item DELETED
\end{itemize}
\end{itemize}
See section \ref{Section: Job Flow} for a description of the various
states. Additionally each value can be prepended the prefix {}``PENDING:''
(like PENDING:ACCEPTED, see section \ref{Section: Job Flow}). This
is used to show that a job is \emph{ready} to be moved to the next
state but it has to stay in it's current state \emph{only} because
otherwise some limits set in the configuration would be exceeded.

\begin{itemize}
\item \textit{job.ID.description} - contains the description of the job
(JD).
\item \textit{job.ID.local} - information about the job used by the A-REX.
It consists of lines of format \textit{{}``name = value''}. Not
all of them are always available. The following names are defined:

\begin{itemize}
\item \textit{subject} - user certificate's subject, also known as the distinguished
name (DN)
\item \textit{starttime} - GMT time when the job was accepted represented
in the Generalized Time format of LDAP 
\item \textit{lifetime} - time period to preserve the SD after the job has
finished in seconds
\item \textit{cleanuptime} - GMT time when the job should be removed from
the cluster and it's SD deleted in Generalized Time format
\item \textit{notify} - email addresses and flags to send mail to about
the job specified status changes
\item \textit{processtime} - GMT time when to start processing the job in
Generalized Time format
\item \textit{exectime} - GMT time when to start job execution in Generalized
Time format
\item \textit{expiretime} - GMT time when the credentials delegated to the
job expire in Generalized Time format
\item \textit{rerun} - number of retries left to rerun the job
\item \textit{jobname} - name of the job as supplied by the user
\item \textit{lrms} - name of the LRMS backend to be used for local submission
\item \textit{queue} - name of the queue to run the job at
\item \textit{localid} - job id in LRMS (appears only after the job has
reached state \textbf{InLRMS})
\item \textit{args} - executable name followed by a list of command-line
arguments
\item \textit{downloads} - number of files to download into the SD before
execution
\item \textit{uploads} - number of files to upload from the SD after execution
\item \textit{gmlog} - directory name which holds files containing information
about the job when accessed through GridFTP interface
\item \textit{clientname} - name (as provided by the user interface) and
IP address:port of the submitting client machine
\item \textit{clientsoftware} - version of software used to submit the job
\item \textit{sessiondir} - the job's SD
\item \textit{failedstate} - state in which job failed (available only if
it is possible to restart the job)
\item \textit{jobreport} - URL of a user requested \emph{logger service}.
The A-REX will also send job records to this service in addition to
the default logger service configured in the configuration (This feature
does not work in current implementation).
\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

\begin{itemize}
\item \textit{job.ID.input} - list of input files. Each line contains 2
values separated by a space. First value contains name of the file
relative to the SD. Second value is an URL or a file description.
Example:

@@

\textit{file description} - {[}size]{[}.checksum].

\hspace*{1cm}\textit{size} - size of the file in bytes.

\hspace*{1cm}\textit{checksum} - checksum of the file identical to
the one produced by \textbf{\textit{cksum}} (1).

Both size and checksum can be left out. Special kind of file description
{*}.{*} is used to specify files which are \textbf{not} required to
exist.

\end{itemize}
This file is used by the '\textbf{\textit{downloader}}' utility. Files
with \emph{URL} will be downloaded to the SD and files with 'file
description' will simply be checked to exist. Each time a new \textbf{valid}
file appears in the SD it is removed from the list and \textit{job.ID.input}
is updated. Any external tool can thus track the process of collecting
input files by checking \textit{job.ID.input.}

\begin{itemize}
\item \textit{job.ID.output} - list of output files. Each line contains
1 or 2 values separated by a space. First value is the name of the
file relative to the SD. The second value, if present, is a URL. Supported
URLs are the same as those supported by job.ID.input.
\end{itemize}
This file is used by the '\textbf{\textit{uploader}}' utility. Files
with \textit{URL} will be uploaded to SE and remaining files will
be left in the SD. Each time a file is uploaded it is removed from
the list and \textit{job.ID.output} is updated. Files not mentioned
as output files are removed from the SD at the the beginning of the
\textbf{Finishing} state.

\begin{itemize}
\item \textit{job.ID.failed} - the existence of this file marks the failure
of the job. It can also contain one or more lines of text describing
the reason of failure. Failure includes the return code different
from zero of the job itself.
\item \textit{job.ID.errors} - this file contains the output produced by
external utilities like \textbf{\textit{downloader}}, \textbf{\textit{uploader}},
script for job submission to LRMS, etc on their stderr handle. Those
are not necessarily errors, but can be just useful information about
actions taken during the job processing. In case of problem include
content of that file while asking for help.
\item \textit{job.ID.diag} - information about resources used during execution
of job and other information suitable for diagnostics and statistics.
It's format is similar to that of \textit{job.ID.local}. The following
names are at least defined:

\begin{itemize}
\item \textit{nodename} - name of computing node which was used to execute
job,
\item \textit{runtimeenvironments} - used runtime environments separated
by ';',
\item \textit{exitcode} - numerical exit code of job,
\item \textit{frontend\_distribution} - name and version of operating system
distribution on frontend computer,
\item \textit{frontend\_system} - name of operating on frontend computer,
\item \textit{frontend\_subject} - subject (DN) of certificate representing
frontend computer,
\item \textit{frontend\_ca} - subject (DN) of issuer of certificate representing
frontend computer,
\end{itemize}
and other information provided by GNU \emph{time} utility. Note that
some implementation of \emph{time} insert unrequested information
in their output. Hence some lines can have broken format.

\item \textit{job.ID.proxy} -delegated X509 credentials.
\item \textit{job.ID.proxy.tmp} - temporary X509 credentials with different
UNIX ownership used by processes run with effective \emph{user id}
different form job owner's \emph{id}.
\end{itemize}
There are other files with names like job.ID.{*} which are created
and used by different parts of the A-REX. Their presence in the \textit{control
directory} can not be guaranteed and can change depending on changes
in the A-REX code.

@@

A-REX Web Service Interface provides means to submit a description
of a computational job to a computing resource, to stage-in additional
data, to monitor and control processing of jobs, and obtain data produced
during the execution of a job. The WS Interface is built and deployed
inside the Hosting Environment Daemon (HED) infrastructure\cite{hed}.

@@

The job submission and control interface is based on a document produced
by the OGF OGSA Basic Execution Services (BES) Working Group\cite{ogsa-bes}.

The exchange of SOAP messages is performed via HTTP(S). The BES interface
is represented by two port-types - BES-Management and BES-Factory.
The former is made to control the A-REX service itself and thus defines
operations to start and stop the functionality of the BES service.
The A-REX does not implement remote control of service functionality.
Hence the BES-Management port-type is not functional. The BES-Factory
port-type provides operations to submit new jobs (to create an activity
in terms of BES) and to monitor its state. It also has an ability
to provide information about the service. A-REX fully implements the
functionality of this port-type.

For job descriptions A-REX accepts the Job Submission Description
Language (JSDL)\cite{jsdl} documents as defined by the OGF Job Submission
Description Language Working Group. Supported elements and extensions
are described below.

@@

A-REX accepts JSDL documents having the following additional elements
(see Appendix E):

>>>>

\documentclass{article}                            %for shorter notes
\usepackage{graphics}                              %for EPS images (latex)
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{listings}                              %for code listings
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\lstloadlanguages{bash,XML}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{array}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {A.Konstantinov},
  pdftitle = {The ARC Computational Job Management Component -- A-REX},
  pdfsubject = {},
  pdfkeywords = {},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
\definecolor{lightblue}{cmyk}{0.05,0,0,0.05}
\lstset{
frame=single,
backgroundcolor=\color{shadecolor},
fillcolor=\color{shadecolor},
rulecolor=\color{lightblue},
basicstyle=\ttfamily,
xleftmargin=0.05\linewidth,
aboveskip=2\medskipamount
}
%
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

@@

%-------- Change this to NORDUGRID-XXXXXXX-NN

@@

%%%%---- The title ----
{\centering \textsc{\Large ARC Computational Job Management Component -- A-REX}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Description and Administrator's Manual}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large A.~Konstantinov\footnote{aleks@fys.uio.no} \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

%\tableofcontents                          %Comment if use article style
\newpage
\section{Introduction\label{sec:intro}}

A-REX is one of ARC middleware components that implements functions
of the so-called \emph{Computing Element} (CE). Here Computing Element
is a service accepting requests containing description of generic
computational job and executing it in the underlying local batch system.

It takes care of job pre- and post-processing. This comprises
stage-in of files containing input data or program modules from wide
range of sources and transfer or storing of output results.

A-REX implements a Web Service (WS) interface which provides a
way to submit and control computational tasks (jobs) to be executed
by A-REX and the underlying Local Resource Management System.

\begin{framed}You should use this document for advanced configuration purposes
and understanding of internals of the aforementioned tools. For general
installation and configuration of whole system refer to other documents
available at \url{http://www.nordugrid.org/papers.html}.
\end{framed}

@@

Each job gets an identifier (\textit{jobid}). This is a handle -- WS-Addressing~\cite{ws-addr-soap} XML document -- which identifies the job in the
A-REX and in the Information Interface.

Jobs are initiated and controlled through the WS interface. Complete
job descriptions (JD) are passed to the A-REX through WS in JSDL~\cite{jsdl}
coded description. Input data files and job executables are transferred
separately through the same interface, as described in Section~\ref{sec:input/output data}.

@@

One of the most important tasks of the A-REX is to take care of processing of the
input and output data (files) of the job. Input files are gathered
in the SD or in the associated cache area. There are two ways to put a file
into the SD:

\begin{itemize}
\item Download is initiated by the A-REX -- This is the case for files defined
in the JD (with name and source). The A-REX alone is responsible to
ensure that all required files will be available in the SD.\\
The supported protocols for sources at the moment are (in case of
full installation): GridFTP, FTP, HTTP, HTTPS (HTTP over SSLv3). Also
some less standard sources are supported. Those are described below.
The A-REX fully relies on HED framework~\cite{hed} for data transferring
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-REX. Examples of such
files are the main executable of the job, the job's input files, etc.
These files can (and should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-REX. Access to AFS, NFS, FTP, HTTP and
any other remote data transport during execution of a job is not guaranteed
(at least not by A-REX).

@@

\section{Job flow\label{Section:Job Flow}}

From the point of view of the A-REX a job passes through various states.
Figure~\ref{job states diagram} presents a diagram of the possible
states of a job.%

\begin{figure}[th]
\centering \includegraphics{pic1.pdf}
\caption{\label{job states diagram}Job states}
\end{figure}

A user can examine the state of a job by querying the dedicated Local
Information Description Interface of A-REX using the UI or any other
suitable tool or through query method of WS interface.

@@

\begin{itemize}
\item \textbf{Accepted} -- In this state the job has been submitted to a
CE but is not processed yet. The A-REX will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} -- The input data is being gathered in the SD (stage-in).
The A-REX is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If all files
are successfully gathered the job moves to the next state. If \textbf{any}
file can't be downloaded or it takes the UI too long to upload a file,
the job moves to \textbf{Finishing} state. It is possible to put a
limit on the number of simultaneous \textbf{Preparing} jobs. If this
limit is exceeded, jobs ready to enter the \textbf{Preparing} state
 will stay in the \textbf{Accepted} state, but prefixed with the PENDING:
mark. Exceptions are jobs which have no files to be downloaded. These
are processed out of limits.
\item \textbf{Submiting} -- The job is being passed for execution to the
\emph{Local Resource Management System} (LRMS). The corresponding
backends for many LRMSs are provided with the default installation.
If the local job submission is successful the job moves to the \textbf{Executing}
state. Otherwise it moves to \textbf{Finishing}. It is possible to
limit the aggregate number of jobs in \textbf{Submiting} and \textbf{Executing}
states.
\item \textbf{Executing} -- The job is queued or being executed in the LRMS.
The A-REX takes no actions except waiting until the job finishes.
\item \textbf{Killing} -- Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} -- The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at an Indexing Service. The user can download data files
from the SD by using the UI or other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} -- No more processing is performed by the A-REX.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the Information Interface of the A-REX. If a
job was moved to \textbf{Finished} because of failure, it may be restarted
on request of a client. When restarted, a job is moved to the state
previous to the one in which it failed and is assigned mark PENDING.
This is needed in order to not break the configuration limits. Exception
is a job failed in \textbf{Executing} state and lacking input files
specified in JD. Such a job is treated like failed in \textbf{Preparing}
state.
\item \textbf{Deleted} -- The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by
the user}. No files will be moved to their destination SE.

\section{URLs\label{sec:urls}}

In full installation the A-REX and it's components support the following
data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
http, https, lfc} and \emph{rls.} For more information please see
{}``The Hosting Environment of the Advanced Resource Connector middleware''
document~\cite{hed}.

\section{Internals\label{section:internals}}

@@

For each local UNIX user listed in the A-REX configuration -- including
a generic one which covers all local user identities -- a \textit{control
directory} exists. In this directory the A-REX stores information
about jobs belonging to that user. Multiple users can share the same
\textit{control directory}. In most common configuration case A-REX
serves all users defined by Operating System and stores their control
file in same directory. To make it easier to recover in case of failure,
the A-REX stores most information in files rather than in memory.
All files belonging to the same job have names starting with \textbf{job.ID.,}
where ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.status} -- current state of the job. This is a plain
text file containing a single word representing the internal name
of current state of the job. Possible values and corresponding external
job states are:

\begin{itemize}
\item ACCEPTED
\item PREPARING
\item SUBMIT
\item INLRMS
\item FINISHING
\item FINISHED
\item CANCELING
\item DELETED
\end{itemize}
\end{itemize}

See Section~\ref{Section:Job Flow} for a description of the various
states. Additionally each value can be prepended the prefix {}``PENDING:''
(like PENDING:ACCEPTED, see Section~\ref{Section:Job Flow}). This
is used to show that a job is \emph{ready} to be moved to the next
state but it has to stay in it's current state \emph{only} because
otherwise some limits set in the configuration would be exceeded.

\begin{itemize}
\item \textit{job.ID.description} -- contains the description of the job
(JD).
\item \textit{job.ID.local} -- information about the job used by the A-REX.
It consists of lines of format \textit{{}``name = value''}. Not
all of them are always available. The following names are defined:

\begin{itemize}
\item \textit{subject} -- user certificate's subject, also known as the distinguished
name (DN)
\item \textit{starttime} -- GMT time when the job was accepted represented
in the Generalized Time format of LDAP 
\item \textit{lifetime} -- time period to preserve the SD after the job has
finished in seconds
\item \textit{cleanuptime} -- GMT time when the job should be removed from
the cluster and it's SD deleted in Generalized Time format
\item \textit{notify} -- email addresses and flags to send mail to about
the job specified status changes
\item \textit{processtime} -- GMT time when to start processing the job in
Generalized Time format
\item \textit{exectime} -- GMT time when to start job execution in Generalized
Time format
\item \textit{expiretime} -- GMT time when the credentials delegated to the
job expire in Generalized Time format
\item \textit{rerun} -- number of retries left to rerun the job
\item \textit{jobname} -- name of the job as supplied by the user
\item \textit{lrms} -- name of the LRMS backend to be used for local submission
\item \textit{queue} -- name of the queue to run the job at
\item \textit{localid} -- job id in LRMS (appears only after the job has
reached state \textbf{InLRMS})
\item \textit{args} -- executable name followed by a list of command-line
arguments
\item \textit{downloads} -- number of files to download into the SD before
execution
\item \textit{uploads} -- number of files to upload from the SD after execution
\item \textit{gmlog} -- directory name which holds files containing information
about the job when accessed through GridFTP interface
\item \textit{clientname} -- name (as provided by the user interface) and
IP address:port of the submitting client machine
\item \textit{clientsoftware} -- version of software used to submit the job
\item \textit{sessiondir} -- the job's SD
\item \textit{failedstate} -- state in which job failed (available only if
it is possible to restart the job)
\item \textit{jobreport} -- URL of a user requested \emph{logger service}.
The A-REX will also send job records to this service in addition to
the default logger service configured in the configuration (This feature
does not work in current implementation).
\end{itemize}
\end{itemize}

This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

\begin{itemize}
\item \textit{job.ID.input} -- list of input files. Each line contains 2
values separated by a space. First value contains name of the file
relative to the SD. Second value is an URL or a file description.
Example:

@@

\textit{file description} -- {[}size]{[}.checksum].

\hspace*{1cm}\textit{size} -- size of the file in bytes.

\hspace*{1cm}\textit{checksum} -- checksum of the file identical to
the one produced by \textbf{\textit{cksum}} (1).

Both size and checksum can be left out. Special kind of file description
{*}.{*} is used to specify files which are \textbf{not} required to
exist.
\end{itemize}

This file is used by the '\textbf{\textit{downloader}}' utility. Files
with \emph{URL} will be downloaded to the SD and files with 'file
description' will simply be checked to exist. Each time a new \textbf{valid}
file appears in the SD it is removed from the list and \textit{job.ID.input}
is updated. Any external tool can thus track the process of collecting
input files by checking \textit{job.ID.input.}

\begin{itemize}
\item \textit{job.ID.output} -- list of output files. Each line contains
1 or 2 values separated by a space. First value is the name of the
file relative to the SD. The second value, if present, is a URL. Supported
URLs are the same as those supported by job.ID.input.
\end{itemize}

This file is used by the '\textbf{\textit{uploader}}' utility. Files
with \textit{URL} will be uploaded to SE and remaining files will
be left in the SD. Each time a file is uploaded it is removed from
the list and \textit{job.ID.output} is updated. Files not mentioned
as output files are removed from the SD at the the beginning of the
\textbf{Finishing} state.

\begin{itemize}
\item \textit{job.ID.failed} -- the existence of this file marks the failure
of the job. It can also contain one or more lines of text describing
the reason of failure. Failure includes the return code different
from zero of the job itself.
\item \textit{job.ID.errors} -- this file contains the output produced by
external utilities like \textbf{\textit{downloader}}, \textbf{\textit{uploader}},
script for job submission to LRMS, etc on their stderr handle. Those
are not necessarily errors, but can be just useful information about
actions taken during the job processing. In case of problem include
content of that file while asking for help.
\item \textit{job.ID.diag} -- information about resources used during execution
of job and other information suitable for diagnostics and statistics.
It's format is similar to that of \textit{job.ID.local}. The following
names are at least defined:
\begin{itemize}
\item \textit{nodename} -- name of computing node which was used to execute
job,
\item \textit{runtimeenvironments} -- used runtime environments separated
by ';',
\item \textit{exitcode} -- numerical exit code of job,
\item \textit{frontend\_distribution} -- name and version of operating system
distribution on frontend computer,
\item \textit{frontend\_system} -- name of operating on frontend computer,
\item \textit{frontend\_subject} -- subject (DN) of certificate representing
frontend computer,
\item \textit{frontend\_ca} -- subject (DN) of issuer of certificate representing
frontend computer,
\end{itemize}
and other information provided by GNU \emph{time} utility. Note that
some implementation of \emph{time} insert unrequested information
in their output. Hence some lines can have broken format.
\item \textit{job.ID.proxy} -delegated X509 credentials.
\item \textit{job.ID.proxy.tmp} -- temporary X509 credentials with different
UNIX ownership used by processes run with effective \emph{user id}
different form job owner's \emph{id}.
\end{itemize}

There are other files with names like job.ID.{*} which are created
and used by different parts of the A-REX. Their presence in the \textit{control
directory} can not be guaranteed and can change depending on changes
in the A-REX code.

@@

A-REX Web Service Interface provides means to submit a description
of a computational job to a computing resource, to stage-in additional
data, to monitor and control processing of jobs, and obtain data produced
during the execution of a job. The WS Interface is built and deployed
inside the Hosting Environment Daemon (HED) infrastructure~\cite{hed}.

@@

The job submission and control interface is based on a document produced
by the OGF OGSA Basic Execution Services (BES) Working Group~\cite{ogsa-bes}.

The exchange of SOAP messages is performed via HTTP(S). The BES interface
is represented by two port-types -- BES-Management and BES-Factory.
The former is made to control the A-REX service itself and thus defines
operations to start and stop the functionality of the BES service.
The A-REX does not implement remote control of service functionality.
Hence the BES-Management port-type is not functional. The BES-Factory
port-type provides operations to submit new jobs (to create an activity
in terms of BES) and to monitor its state. It also has an ability
to provide information about the service. A-REX fully implements the
functionality of this port-type.

For job descriptions A-REX accepts the Job Submission Description
Language (JSDL)~\cite{jsdl} documents as defined by the OGF Job Submission
Description Language Working Group. Supported elements and extensions
are described below.

@@

A-REX accepts JSDL documents having the following additional elements
(see Appendix~\ref{annex:jsdl-extension}):
